Timing and Likelihood of the Origin of Life Derived from Post-Impact Highly Reducing Atmospheres

### 2023-09-12 ms.tex
tiny tweaks to make kevin happy

\begin{abstract}
Big impacts on the early Earth would have created highly reducing atmospheres that generated molecules needed for the origin of life, such as nitriles. However, such impactors can be followed by collisions that were still sufficiently big to vaporize the ocean and destroy any pre-existing life. Thus, a post-impact reducing atmosphere that gives rise to life needs to be followed by a lack of subsequent sterilizing impacts for life to persist. We assume that prebiotic chemistry required a post-impact reducing atmosphere. Then, using statistics for the impact history on early Earth and the minimum impact mass needed to generate post-impact highly reducing atmospheres, we show that the median timing of impact-driven biopoiesis (i.e., the origin of life) is favored early in the Hadean eon. However, uncertainties are large because impact bombardment is stochastic, so biopoiesis could have occurred in a $\sim 0.5$ billion year window from 4.45 to 3.9 Ga within 95\% uncertainty. Previous claims of a far narrower window for the origin of life are unsupported. In an optimistic scenario for life starting from post-impact reducing atmospheres, we find that the origin of life is possible in $\sim 90\%$ of stochastic impact realizations. In our most pessimistic case, life's origin is still fairly likely ($\sim 20\%$ chance). This potentially bodes well for biopoesis on rocky planets elsewhere which will have experienced an early episode of enhanced impact bombardment given how planets form.
\end{abstract}

@@

Figure \ref{fig:example_impact_histories}c shows an impact history where Moneta does not occur. Instead, Earth's mantle HSEs are delivered by multiple $10^{21}$ to $4 \times 10^{21}$ kg collisions. Whether this bombardment history would cause biopoiesis depends on the required minimum impact mass to produce significant origin-of-life nitriles (e.g. HCN and HCCCN). \citet{Wogan_2023} used photochemical models of post-impact atmospheres to show that, optimistically, an impact $> 4 \times 10^{20}$ kg should give rise to an atmosphere that makes significant HCN and HCCCN. Under this optimistic requirement, the last ocean-vaporizing impact in Figure \ref{fig:example_impact_histories}c could successfully start life and occurs at 4.38 Ga with mass $3 \times 10^{21}$ kg. However, with pessimistic modeling assumptions, \citet{Wogan_2023} found that a $> 5 \times 10^{21}$ kg impactor is instead required for substantial nitrile delivery to the surface. In this alternative scenario, an origin of life never occurs in Figure \ref{fig:example_impact_histories}c because there is no impactor larger than $5 \times 10^{21}$ kg.

>>>>

\begin{abstract}
Big impacts on the early Earth would have created highly reducing atmospheres that generated molecules needed for the origin of life, such as nitriles. However, such impactors can be followed by collisions that were still sufficiently big to vaporize the ocean and destroy any pre-existing life. Thus, a post-impact reducing atmosphere that gives rise to life needs to be followed by a lack of subsequent sterilizing impacts for life to persist. We assume that prebiotic chemistry required a post-impact reducing atmosphere. Then, using statistics for the impact history on early Earth and the minimum impact mass needed to generate post-impact highly reducing atmospheres, we show that the median timing of impact-driven biopoiesis (i.e., the origin of life) is favored early in the Hadean eon. However, uncertainties are large because impact bombardment is stochastic, so biopoiesis could have occurred in a $\sim 0.5$ billion year window from 4.45 to 3.9 Ga within 95\% uncertainty. Previous claims of a far narrower window for the origin of life are unsupported. In an optimistic scenario for life starting from post-impact reducing atmospheres, we find that the origin of life is favorable in $\sim 90\%$ of stochastic impact realizations. In our most pessimistic case, life's origin is still fairly likely ($\sim 20\%$ chance). This potentially bodes well for biopoiesis on rocky planets elsewhere which will have experienced an early episode of enhanced impact bombardment given how planets form.
\end{abstract}

@@

Figure \ref{fig:example_impact_histories}c shows an impact history where Moneta does not occur. Instead, Earth's mantle HSEs are delivered by multiple $10^{21}$ to $4 \times 10^{21}$ kg collisions. Whether this bombardment history would favor biopoiesis depends on the required minimum impact mass to produce significant origin-of-life nitriles (e.g. HCN and HCCCN). \citet{Wogan_2023} used photochemical models of post-impact atmospheres to show that, optimistically, an impact $> 4 \times 10^{20}$ kg should give rise to an atmosphere that makes significant HCN and HCCCN. Under this optimistic requirement, the last ocean-vaporizing impact in Figure \ref{fig:example_impact_histories}c could successfully start life and occurs at 4.38 Ga with mass $3 \times 10^{21}$ kg. However, with pessimistic modeling assumptions, \citet{Wogan_2023} found that a $> 5 \times 10^{21}$ kg impactor is instead required for substantial nitrile delivery to the surface. In this alternative scenario, an environment favorable for an origin of life never occurs in Figure \ref{fig:example_impact_histories}c because there is no impactor larger than $5 \times 10^{21}$ kg.
